% Sebastian Daza - Curriculum Vitae

\documentclass[10pt,letterpaper]{article}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{mathptmx}

% Fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{fontspec}
%\usepackage{mathpazo}
% \usepackage[urw-garamond]{mathdesign}
\usepackage{ebgaramond}
%\usepackage{fourier}

% Set your name here
\def\name{Sebastian Daza}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {sociology, demography, quantitative methods, criminology},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 9.0in},
  left=1.0in,
  top=1.0in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets and compact spacing
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
    \setlength{\itemsep}{0.25em}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0.25em}
  }
}{
  \end{list}
}


\begin{document}

% Place name at left
%{\huge \name}

% Alternatively, print name centered and bold:
\centerline{\huge \name}

\bigskip
\bigskip

\begin{minipage}[t]{.75\textwidth}
  Department of Sociology \\
  University of Wisconsin-Madison \\
  1180 Observatory Drive \\
  Madison, WI 53706-1393 \\
\end{minipage}
\begin{minipage}[t]{0.3\textwidth}
  % Phone: (608) 628-8548 \\
  Email: \href{mailto:sdaza@ssc.wisc.edu}{sdaza@ssc.wisc.edu} \\
  \href{https://sdaza.com/}{https://sdaza.com} \\

\end{minipage}

\bigskip
\bigskip

\section*{Education}

\begin{itemize}
  \item Ph.D., Sociology - Demography, University Wisconsin--Madison, 2021.
  \item Fellow, \href{https://www.thedataincubator.com/fellowship.html}{The Data Incubator}, 2017.
    % Received May 16, 2010
  \item M.A. Sociology, Catholic University of Chile, 2009.
    % Received May 14, 2006
  \item B.S. Sociology, Catholic University of Chile, 2004.
    % Received May 14, 2004
\end{itemize}

\section*{Fields of Research Interest}

\begin{itemize}
\item Social Demography, Health, Social Complexity, Social Networks, Deviance and Crime, Surveys and Methods.
\end{itemize}


\section*{Research}

\subsection*{Peer-Reviewed Articles}

\begin{itemize}

\item Daza, Sebastian, Alberto Palloni and Jerrett Jones (2020).
  \textit{\href{https://rdcu.be/b278V}
  {The Consequences of Incarceration for Mortality for the United States}
  }.
  Demography.

\item Venkataramani, Atheendar S., Sebastian Daza and Ezekiel Emanuel (2020).
  \textit{
  \href{doi:10.1001/jamainternmed.2019.6532}
  {Association of Social Mobility With the Income-Related Longevity Gap in the United States: A Cross-Sectional, County-Level Study}
  }.
 JAMA Internal Medicine.


\item Larroulet, Pilar, Droppelmann, Catalina, Del Villar, Paloma, Daza, Sebastian, Figueroa, Ana, and Osorio, Victoria (2020).
    \textit{
  \href{https://www.crimejusticejournal.com/article/view/1467}
  {Who Is Transitioning out of Prison? Characterising Female Offenders and Their Needs in Chile}
  }.
International Journal for Crime, Justice and Social Democracy.

 \item Daza, Sebastian and Kurt Kreuger (2019).
  \textit{
  \href{https://doi.org/10.1177/0049124119826147}
  {Agent-Based Models for Assessing Complex Statistical Models: An Example Evaluating Selection and Social Influence Estimates from the SIENA model}
  }.
 Sociological Methods and Research.


\end{itemize}

\subsection*{Work in progress}

\begin{itemize}

\item Daza, Sebastian and Alberto Palloni (2021). {\textit{Early Exposure to County Income Mobility and Adult Individual Health in the United States.}}
\item Larroulet, Pilar, Sebastian Daza and Ignacio Borquez (2021). {\textit{Employment Trajectories among Females Returning from Prison.}}
\item Voigt, Mathias, Daza, Sebastian, Ordanovich, Dariya, and Palloni, Alberto (2021). {\textit{Trends in Education-Specific Differences in Disability-Free Life Expectancy in Spain, 2008-2017}}.
\item Daza, Sebastian and Alberto Palloni (2019). {\textit{Income Inequality, Income Mobility, and Mortality in the U.S.}} Under review.
\item Palloni, Alberto, Hiram Beltrán-Sánchezx and Sebastian Daza (2021). {\textit{Uncertainty and Mortality Estimates in the Latin American and Caribbean (LAC) Region.}}

\end{itemize}

\subsection*{Other Publications and Reports}

\begin{itemize}

\item Somma, Nicolas and Sebastian Daza (2011).
  \textit{
  \href{https://sdaza.com/assets/pdf/articulo-nicolas-somma.pdf}
  {Why did the Concertación Lose? Age, Egotropic Vote, and Political Divisions in the 2009--2010 Chilean Election}
  }.
  Cuadernos ISUC, Vol. 1, Num. 3.

\item Daza, Sebastian (2009). Complementary Study on the Evaluation of the 2003--2008, National Strategy Against Drug Use. National Budget Agency for the Exchequer, Chile.

\item Daza, Sebastian, Daniella Leal and Von Hausen, Carolina (2009). Design and Validation of Context Questionnaires for the National System of Educational Assessment. Ministry of Education, Chile.

\item Cereceda, Luz, Sebastian Daza and Daniella Leal (2008). Results of the National Survey \textit{Actors in the Educational System}. Ministry of Education, Chile.

\item Valenzuela, Eduardo, Sebastian Daza and Andrés Biehl (2007). Results of the Survey \textit{Social Cohesion in Latin America}. Corporation for Latin American Studies, Chile.

\item Valenzuela, Eduardo and Sebastian Daza (2006). A Study on Drug Use and Associated Factors Among Underage Offenders. National Service of Minors, Chile.

\item Herrera, Soledad, Sebastian Daza, Daniella Leal and Eliana Vivencio (2006). A Series of Studies on Socio--Economic Characterization, 2003. Ministry of Planning and Cooperation, Chile.

\item Daza, Sebastian and Darinka Radovic (2005). Report on the Evaluation of the Unit of School Psycho--Social Care and Welfare. Corporación Municipal Puente Alto, Chile.

\item Daza, Sebastian and  Macarena Vivent (2004). Report on the Evaluation of the Project SaludArte. Corporación Municipal Puente Alto, Chile.

\item Daza, Sebastian and Macarena Vivent (2003). Towards the Creation of a Quality-of-Life Instrument for Minors Between Ages 0 and 3. Department of Public Health, Catholic University of Chile.


\end{itemize}

\subsection*{Unpublished Working Papers}

\begin{itemize}

\item Palloni, Alberto and Sebastian Daza (2014).
   \textit{Intergenerational Transfers in the U.S., Economic Inequality, and Social Stratification.}

\item Daza, Sebastian and Jason Fletcher (2016).
  \textit{Consequences of Childbearing in Delinquency and Substance Use.}

\item Daza, Sebastian, Eric Grodsky and Ann Meier (2014).
  \textit{Exploring the Consequences of Adolescent Sex in Behavioral Context.}

\item Daza, Sebastian (2013).
  %\href{}
  \textit{Drug Use and Crime Transitions Among Young Offenders in Chile}.
  % Revise and resubmit at \textit{Econometrics Journal}.

\item Daza, Sebastian (2013).
  %\href{}
  \textit{Segregation and Student Composition Effects on Academic Achievement in Chile}.
  %Revise and resubmit at \textit{Econometrics Journal}.

\item Daza, Sebastian (2012).
  %\href{}
  \textit{Substance Use, Family, and Peers Effects among Chilean Adolescents}.
  %Revise and resubmit at \textit{Quantitative Economics}.

\item Daza, Sebastian (2011).
  %\href{}
  \textit{Patterns and Progressions of Substance Use among Chilean Adolescents}.
  %Revise and resubmit at \textit{Quantitative Economics}.

\end{itemize}

\section*{Professional Experience}

\subsection*{\href{http://iegd.csic.es/}{Institute of Economy Geography and Demography, CSIC-CCHS, Spain}}

\begin{itemize}
\item \href{https://echo-erc.csic.es/}{Senior Researcher ECHO, European Research Council (ERC) project, December 2019--Present.}
\end{itemize}

\subsection*{\href{https://juntosglobal.com/}{Juntos Global, USA}}

\begin{itemize}
\item \href{https://juntosglobal.com/}{Senior Data Scientist, May 2018--December 2019.}
\end{itemize}

\subsection*{\href{http://sociologia.uc.cl/}{University of Wisconsin-Madison, US}}

\begin{itemize}
\item \href{http://www.ssc.wisc.edu/cde/people/grads.php}{Research Assistant, Center for Demography and Ecology UW--Madison, September 2014--August 2018.}
\item \href{http://www.ssc.wisc.edu/cde/people/grads.php}{Project Assistant, Center for Demography and Ecology UW--Madison, January 2011--August 2014.}
\end{itemize}

\subsection*{\href{http://sociologia.uc.cl/}{Catholic University of Chile}}
 \begin{itemize}
 \item \href{http://justiciaysociedad.uc.cl/nosotros/investigadores-asociados}{Research Associate, Center for Justice and Society Studies, August 2016--present.}
 \item \href{http://sociologia.uc.cl/desuc}{Research Associate, Department of Sociology, August 2010--December 2017}
 \item Director of Applied Studies (DESUC), Department of Sociology, 2005--2010.
 \item Lecturer, Department of Sociology, 2008--2010.
 \item Instructor, Department of Sociology, \href{http://www.educacioncontinua.uc.cl}{Professional Certificate in Quantitative Methods}, 2008--2010.
 \end{itemize}

\subsection*{\href{http://www.udp.cl/facultades_carreras/ciencias_sociales-historia/}{Diego Portales University, Chile}}

\begin{itemize}
\item Lecturer, Department of Sociology, 2007.
\end{itemize}

\subsection*{Non-Academic Institutions}

\begin{itemize}
\item \href{http://juntosglobal.com/}{Senior Quantitative Social Scientitst, Juntos Inc, US, May 2018--present.}
\item Principal Researcher, Evaluation of Program \textit{School Psycho-Social Care and Welfare}, Corporación Municipal de Puente Alto, Chile, 2003--2008.
\item \href{http://www.latinobarometro.org}{Researcher, Corporación Latinobarómetro, Chile, 2005.}
\item \href{http://www.feedback.cl}{Researcher, Feedback Consultancy, Chile, 2004.}
\end{itemize}


\section*{Teaching}

\subsection*{\href{http://sociologia.uc.cl/}{Catholic University, Department of Sociology}}

\begin{itemize}
\item Advanced Methods of Social Research, 2008--2009.
\item Qualitative Analysis Techniques using \href{http://www.qsrinternational.com/}{Nvivo}, 2009.
\item Advanced Analysis of Social Data, 2008--2009.
\item An Introduction to the Analysis of Social Data, 2009.
\item Design and Implementation of Social Surveys, 2008.
\end{itemize}

\subsection*{\href{http://www.udp.cl/facultades_carreras/ciencias_sociales-historia/}{Diego Portales University, Chile, Department of Sociology}}

\begin{itemize}
\item Introduction to Statistics, 2007.
\end{itemize}


\section*{Conference and Seminar Presentations}

\begin{itemize}

\item TRENDS Conference: Why is Health in the US Continuing to Lag Behind? May 2021. {\textit{Early Exposure to County Income Mobility and Adult Individual Health in the United States.}}

\item Virtual Annual Meeting of the Population Association of America, May 2021. {\textit{Early Exposure to County Income Mobility and Adult Individual Health in the United States.}}

\item Congreso Internacional de la Sociedad Chilena de Criminología, November 2020. {\textit{Trayectorias Laborales de Mujeres que Salen en
Libertad.}}

\item European Population Conference, 2020. {\textit{Contextual Income Mobility and Individual Health.}}

\item Life Course Influences on Later Life Health and Wellbeing, Institute for Social Research, Ann Arbor, MI, June 2017. {\textit{The Consequences of Incarceration for Mortality.}}


\item Life Course Influences on Later Life Health and Wellbeing, Institute for Social Research, Ann Arbor, MI, June 2017. {\textit{The Consequences of Incarceration for Mortality.}}


\item  Annual Meeting of the Population Association of America, Chicago, IL, April 2017. {\textit{Income Inequality, Income Mobility, and Mortality in the U.S.}}

\item  Annual Meeting of the Population Association of America, Chicago, IL, April 2017. {\textit{Examining the Stochastic Actor-Oriented Model (Siena) for Selection and Influence Estimation under Misspecification}}

\item  Annual Meeting of the Population Association of America, Chicago, IL, April 2017. {\textit{The Consequences of Incarceration for Mortality.}}

\item  Annual Meeting of the Population Association of America, Washington, D.C., March-April 2016. {\textit{Consequences of Childbearing in Delinquency and Substance Use.}}

\item  Annual Meeting of the Population Association of America, Boston, MA, May 2014. {\textit{Private Intergenerational Transfers between Parents and Children in the U.S., and Consequences of the Great Recession.}}

\item  Annual Meeting of the Population Association of America, Boston, MA, May 2014. {\textit{Exploring the Consequences of Adolescent Sex in Behavioral Context.}}

\item Weighting Workshop, University of Wisconsin–Madison, September 2012. \textit{Longitudinal Study of Drug Use in Chile (LSDU): Some Questions about Sampling Weights}.

\item	Presentation to the Milenio Scientific Committee of the results of project Relationship between the Onset of Drug Use and the Initiation in Delinquent Behavior in a Prison Population, 2009

\item Presentation of the results of the Survey on Social Cohesion in Latin America. Methodological Design of the Study and Results on Educational and Social Mobility. \href{http://www.cieplan.org/}{Corporation for Latin American Studies}, 2007.

\item Presentation of the Results of the Program for School Psycho-Social Welfare. Corporación Municipal de Puente Alto, 2005.

\item Chile’s 30th Seminar on Public Health, Diego Portales Convention Centre, October 1st, 2nd and 3rd, 2003. Presentation of the results of the research: Towards the Creation of a Quality-of-Life Instrument for Minors between Ages 0 and 3, 2003.

\end{itemize}

\section*{Professional Activities}

\begin{itemize}
\item Member, Population Association of America, 2012--Present.
\item Mentor for new graduate students, Department of Sociology, UW-Madison,
August 2013 – 2014.
\item Reviewer: \textit{Journal of the American Medical Association (JAMA)}, \textit{PLOS One}, \textit{Social Science and Medicine}, \textit{Social Forces}, \textit{Race and Social Problems}, \textit{Journal of Policy Analysis and Management}.
\end{itemize}

\section*{Grants}

\begin{itemize}
\item National Institute on Aging, Small Grants for Research Using PSID Data, \textit{The Consequences of Incarceration for Health and Mortality},  \$20,000, May 2016 - June 2017
\item Exetension National Institute on Aging, Small Grants for Research Using PSID Data, \textit{The Consequences of Incarceration for Health and Mortality},  \$15,000, May 2017 - June 2018
\item San Carlos de Maipo Foundation, \textit{Chile Re-entry Study}, \$150,000, Jan 2016 - June 2018.
\end{itemize}


\section*{Honors, Awards, \& Fellowships}

\begin{itemize}
\item Grant to attend Agent-based Modeling workshop, Technische Universität Dresden, Germany, 2015.
\item Grant to attend Agent-based Modeling workshop, Max Planck Institute for Demographic Research (MPIDR), Germany, 2014.
\item Best Master's Thesis Award, Catholic University of Chile, 2010.
\item \href{http://www.becaschile.cl}{Chilean Ph.D. Scholarship, BecasChile, 2010-2014.}
\item Scholarship for Master’s Degree at Catholic University of Chile.
\item Scholarship for Undergraduate Studies in Sociology at Catholic University of Chile.
\end{itemize}

\section*{Workshops Attended}

\begin{itemize}


\item Agent-Based and Hybrid Modeling Bootcamp and Incubator for Health Researchers, August 2016. Instructor: Nathaniel Osgood.
\item Agent-based Model Analysis, Technische Universität Dresden, Germany, July 2015. Competitive selection.
\item Agent-based Modeling and Simulation, Max Planck Institute for Demographic Research (MPIDR), Rostock, Germany, October 2014. Competitive selection.
\item The Pathways to Desistance Study: Analyzing the Life Event Calendar Data for Substance Abuse Research. ICPSR/NAHDAP Summer Program Workshop, July 2014, Ann Arbor, MI.
\item Doing Bayesian Data Analysis, University of Wisconsin--Madison, July 2012. Instructor: John Kruschke.
\item Social Network Analysis with Stochastic Actor-based Models, Summer Institutes in Statistics,  University of Kansas, June 2011. Instructor: \href{http://www.ppsw.rug.nl/~steglich/sites/index.htm}{Christian E. G. Steglich}.

 \end{itemize}

\section*{Languages}

\begin{itemize}
\item English (Proficient), Spanish (Native).
\end{itemize}

\section*{Computer Skills}

\begin{itemize}
\item \textit{Statistical Analysis Software:} R, Stata, Mplus, HLM, SPSS, SAS, and CTM.
\item \textit{Data Science}: Machine Learning, NLP, Model Ensemble, Apache Spark, Databricks, Dash, Shiny.
\item \textit{Agent-based Modeling}: Anylogic, NetLogo.
\item \textit{General Programming}: Python and Java.
\item \textit{Social Network Analysis:} R (RSiena, Statnet), Python (NetworkX, graph-tool), Ucinet.
\item \textit{Qualitative Data Analysis Software:} \href{http://www.qsrinternational.com/}{Nvivo}.
\item \textit{Census and Survey Processing:} CsPro.
% \item \textit{Other:} \LaTeX{}.
\end{itemize}

\section*{Software}

\begin{itemize}
\item \href{https://github.com/sdaza/acsr}{\texttt{acsr}: R package to manage data from the US Census American Community Survey (ACS)}. Created for the Applied Population Laboratory at UW--Madison, May 2015.
\item \href{https://github.com/sdaza/sampler}{\texttt{sampler}: R package to compute sample sizes and errors for survey design}.
\end{itemize}

\section*{References}

\begin{itemize}
\item Alberto Palloni, Ph.D.\\
Samuel Preston Professor of Sociology, UW-Madison\\
(608) 262--2182, \href{mailto:palloni@ssc.wisc.edu}{palloni@ssc.wisc.edu}

% \item Jason Fletcher, Ph.D.\\
% Associate Professor of Public Affairs and Sociology, UW-Madison\\
% (608) 262-4436 \href{mailto:jfletcher@ssc.wisc.edu}{jfletcher@ssc.wisc.edu}
% \end{itemize}
\item Jenna Nobles, Ph.D.\\
Assistant Professor of Sociology, UW-Madison\\
(608) 262--4024, \href{mailto:jnobles@ssc.wisc.edu}{jnobles@ssc.wisc.edu}
\end{itemize}

\bigskip
\bigskip
\begin{center}
  \begin{small}
    \textit{Last updated: \today}
  \end{small}
\end{center}

% Footer


\end{document}
