---
layout: post
title: "My Facebook Network"
description: ""
category: SNA
tags: [igraph, home]

---


I am taking a [coursera](https://www.coursera.org) class on Social Network Analysis. The first assignment consisted of plotting your Facebook network. In order to get your Facebook Network you can use the [Netvizz](https://apps.facebook.com/netvizz/) app. This application allows you to create gdf files (a simple text format that specifies an undirected graph) from the friendship relations. I used this [simplified Netvizz version](http://snacourse.com/getnet/) from [coursera](https://www.coursera.org)  that generates gml files directly.


```{r, echo=FALSE, warning=FALSE}
options(replace.assign=TRUE, width=60,  scipen = 4, digits = 3)
graph.basic.stats <- function(g) {
  # check g object
  if (!is.igraph(g)) stop("given grapg is not an igraph object")

  # compute some stats
  n <- vcount(g)                                  # number of nodes
  e <- ecount(g)      					  # number of edges
  deg <- degree(g)                                # degrees
  isol <- sum(deg==0)                             # number of isolated nodes
  comps <- clusters(g, mode="strong")               # Calculate the maximal (weakly or strongly) connected components of a graph

  # component sizes
  compsize <- sort(comps$csize, decreasing=TRUE)
  c1 <- compsize[1]  						# get the biggest component size
  c2 <- compsize[2]							# get the next biggest component size

  # print information
  cat("Number of nodes:", n, "\n")
  cat("Number of edges:", e, "\n\n")

  cat("Degree\n")
  cat("  Average:", mean(deg), "\n\n")

  cat("\n")
  cat("Giant component  Size:", c1, "\n")
  cat("Giant component  As perc.:", c1/n, "\n")
  cat("Second Giant component:", c2, "\n")
  cat("Second Giant component As perc.:", c2/n, "\n\n")

  cat("Isolated nodes\n")
  cat("  Number:", isol, "\n")
  cat("  As perc.:", isol/n, "\n\n")
}

```


```{r}
library(igraph)

# read the facebook network from a gml file
f  <-  read.graph("sdaza.gml",format="gml")
```

In order to plot my Facebook network, I extract the following attributes: gender, wall post counts, and interface language. I also assign some colors and shapes to those attributes.

```{r}
# gender, assigning shapes
ssex  <- V(f)$sex
table(ssex)
ssex[ssex=="female"]  <-  "circle"
ssex[ssex=="male"]  <- "square"
table(ssex)

# interface language, assigning colors
cloc  <- V(f)$locale
table(cloc)

cloc[cloc %in% c("es_ES", "it_IT", "nl_NL", "en_GB", "gl_ES", "ko_KR")]  <-  "Yellow"
cloc[cloc %in% c("es_CL", "pt_BR","es_LA")]  <-  "Green"
cloc[cloc==c("en_US")]  <-  "Blue"
table(cloc)

# wall posts count, assigning node size
nsize  <- (V(f)$wallcount/ max(V(f)$wallcount) + .1 )* 15
```

Now, I can plot the network:

```{r fig1}
plot(f, layout=layout.fruchterman.reingold, edge.arrow.size=.5, vertex.label=NA, vertex.size=3, vertex.color=cloc, vertex.shape=ssex)
```

If we use __wall post counts__ to size the nodes:

```{r fig2}
plot(f, layout=layout.fruchterman.reingold, edge.arrow.size=.5, vertex.label=NA, vertex.size=nsize, vertex.color=cloc, vertex.shape=ssex)
```

We can also obtain some basic descriptive statistics of the network using a __graph.basic.stats__ function (see [here](http://www.isk.kth.se/~shahabm/WSAnalysis/networks/NetworkAnalysis.r)):

```{r}
graph.basic.stats(f)
```

As can be seen, I have relatively compact and homogeneous components based on language interface. You can see how your own Facebook network looks like.