---
layout: post
title: "Simple R package to define sample sizes and MOEs"
description: ""
category: survey
tags: [sampling, home]
---

In this post I present a simple R package called [**sampler**](https://github.com/sdaza/sampler). The package defines **sample sizes** and **margins of error (MOE)** for a proportion, as usually is done when designing public opinion surveys. [In a previous post](http://sdaza.com/survey/2014/01/19/samplesize/), I presented some functions that do mostly the same thing. This new package, however, includes some new features that might be useful.

## Installation

```{r, eval = FALSE}
# you have to install devtools first
devtools::install_github("sdaza/sampler")
library(sampler)
```

## Functions

The packages contains four functions:

- **ssize**: computes sample size.
- **serr**: computes MOE.
- **astrata**: assigns sample sizes to strata.
- **serrst**: computes MOE for stratified samples.



```{r, echo = FALSE}
library(sampler)
```

## Define sample size: *ssize*

```{r}
ssize(.05)
# design effect (deff) and response rate (rr)
ssize(.05, deff = 1.2, rr = .90)
# finite population correction
ssize(.05, deff = 1.2, rr = .90, N = 1000)
# warning message
ssize(.05, deff = 1.2, rr = .90, N = 100)
```

## Define sampling error: *serr*

```{r}
serr(384)
serr(512, deff = 1.2, rr = .90)
serr(370, deff = 1.2, rr = .90, N = 1000)
# we still get an answer
serr(100, deff = 1.2, rr = .90, N = 100)
```

## Strata allocation: *astrata*

These examples show how to allocate a sample size across strata. Look at *?astrata* in **R** for definitions of the allocation procedures that are available.

```{r}
# I will use data.table
library(data.table)
chile <- data.table(chile)
chile
```

```{r}
# proportional for a sample of 1000
chile[, aprop := astrata(1000, wp = 1, N = pob)]

# fixed (same number by stratum)
chile[, afixed := astrata(1000, wp = 0, N = pob)]

# 40% proportional, 60% fixed
chile[, a40 := astrata(1000, wp =.4, N = pob)]

# 60% proportional, 40% fixed
chile[, a60 := astrata(1000, wp =.6, N = pob)]

# square-root
chile[, aroot := astrata(1000, method = "root", N = pob)]

# neyman
chile[, aneyman := astrata(1000, method = "neyman", N = pob, p = pr)]

# standard deviation
chile[, astdev := astrata(1000, method = "stdev", N = pob, p = pr)]

# error
chile[, aerr := astrata(e = .11, method = "error", N = pob, p = pr)]
```

```{r, echo = FALSE}
chile
```

## Getting sampling error from a stratified sample: *serrst*

```{r}
# the second most efficient allocation
serrst(n = chile$aprop, N = chile$pob, p = chile$pr)
# the worst solution
serrst(n = chile$afixed, N = chile$pob, p = chile$pr)
serrst(n = chile$a40, N = chile$pob, p = chile$pr)
serrst(n = chile$a60, N = chile$pob, p = chile$pr)
serrst(n = chile$aroot, N = chile$pob, p = chile$pr)
# the most efficient allocation
serrst(n = chile$aneyman, N = chile$pob, p = chile$pr)
serrst(n = chile$astdev, N = chile$pob, p = chile$pr)
serrst(n = chile$aerr, N = chile$pob, p = chile$pr)
```

## Combining criteria

```{r}
# get error for 60% proportional / 40% fixed allocation for each strata
chile[, error_a60 := serr(a60, p = pr)]

# assign sample sizes assuming 13% error for each strata
chile[, serr13 := astrata(e = .13, method = "error", N = pob, p = pr)]

# total error, not that good!
serrst(n = chile$serr13, N = chile$pob, p = chile$pr)
chile[, .(reg, pob, pr, a60, error_a60, serr13)]
```

We can adjust a bit more:

```{r}
# when error is higher than .13, use serr13
chile[, sfinal := ifelse(error_a60 > .13, serr13, a60)]

# new error by stratum
chile[, error_sfinal := serr(sfinal, p = pr)]

# total error, much better!
serrst(n = chile$sfinal, N = chile$pob, p = chile$pr)

# although the total sample size is now bigger
sum(chile$sfinal)
```
```{r, echo = FALSE}
# details
chile[, .(reg, pob, pr, sfinal, error_sfinal)]
```

That's it. A simple package to do simple calculations.

