---
layout: post
title: "R package to compute statistics from the American Community Survey"
description: ""
category: "R"
tags: [ACS, R]
---


The `acsr` package allows to easily extract and compute statistics from the America Community Survey, and it was created for the Applied Population Laboratory (APT) at  University of Wisconsin-Madison.

<h2 class="section-heading">Installation</h2>

The functions depends on the `acs` and `data.table`, so it is a good idea to install those packages before. The `acsr` package is hosted on a github repository and we will  use the `devtools` library to install the package:

```{r, eval = FALSE}
devtools::install_github("sdaza/acsr")
library(acsr)
```

<h2 class="section-heading">Getting variables</h2>


<h2 class="section-heading">Computing statistics</h2>



## Install `acsr` package in `R`


```{r, eval = FALSE}
# set a server
options(repos=structure(c(CRAN="http://rweb.quant.ku.edu/cran/")))

# install devtoos and acs from CRAN
install.packages("devtools")
install.packages("acs")

# install data.table from github
devtools::install_github("Rdatatable/data.table", build_vignettes = FALSE)

# install acsr from github
devtools::install_github("sdaza/acsr")
```

---

## Load `acsr` and help files

- Remember to set the API key for the `acs` package. You have to do it only once

    ```{r, eval = FALSE}
    api.key.install(key="")
    ```

- Explore help files of the main functions

    ```{r, eval = FALSE}
    library(acsr)
    ?acsdata
    ?sumacs
    ```

- Check defaults!
    ```
    state = WI
    year  = 2013
    span = 5
    level = county
    ```

---

## Why to use the `acsdata` function?

- To download the data can be slow, especially when there are many geographic or administrative levels
- A better approach is to download the data first, and then use them as input

    ```{r, echo=FALSE }
    library(acsr)
    ```

    ```{r }
    d <- acsdata(formula = "(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
     level = "county")
    ```

- This will create a list of ACS objects

---

## The `sumacs` function

```{r }
tab <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
varname = "langspan0913", method = "prop", data = d)
tab[1:3, c(1:4,11:12), with = FALSE]
```

---

## Some options: `format.out`

```{r }
tab <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
varname = "langspan0913", method = "prop", data = d, format.out = "long")
tab[1:3, c(1:4,11:13), with = FALSE]
```

---

## Some options: `file`

- Define a CSV file to export the output

    ```{r }
    tab <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "langspan0913", method = "prop", data = d, file = "out.csv")
    ```

---

## Some options: standard errors and `one.zero`

```{r, warning=FALSE, message=FALSE, results="hide"}
tab1 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
varname = "langspan0913", method = "prop", level = "tract", county = 1,
tract = 950501, one.zero = FALSE)
```

```{r, echo = FALSE}
  tab1[, c(1,6, 11:12), with = FALSE]
```

```{r, warning=FALSE, message=FALSE,  results="hide"}
tab2 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
varname = "langspan0913", method = "prop", level = "tract", county = 1,
tract = 950501, one.zero = TRUE)
```

```{r, echo = FALSE}
tab2[, c(1,6, 11:12), with = FALSE]
```

---

## Some options: `one.zero`

- `one.zero = FALSE`

    ```{r, eval = TRUE}
    ( sqrt(( 5.4711125 ^ 2 + 22.4924 ^ 2 + 5.4711125 ^ 2 ) -
    ( 0.0226390685640362 ^ 2 * 102.1277 ^ 2 )) / 1546 ) * 1.645
    ```

- `one.zero = TRUE`

    ```{r, eval = TRUE}
    ( sqrt(( 5.4711125 ^ 2 + 22.4924 ^ 2) -
    ( 0.0226390685640362 ^ 2 * 102.1277 ^ 2 )) / 1546 ) * 1.645
    ```

---

## `ratio` and `one.zero` I

- `method = "ratio"` and `one.zero = FALSE`

    ```{r, warning=FALSE, message=FALSE,  results="hide"}
    tab3 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
    varname = "langspan0913", method = "ratio", level = "tract", county = 1,
    tract = 950501, one.zero = FALSE)
    ```

    ```{r, echo = FALSE}
      tab3[, c(1,6, 11:12), with = FALSE]
    ```

    ```{r, eval = TRUE}
    ( sqrt(( 5.4711125 ^ 2 + 22.4924 ^ 2 + 5.4711125 ^ 2 )
    + ( 0.0226390685640362 ^ 2 * 102.1277 ^ 2 )) / 1546 ) * 1.645
    ```

---

## `ratio` and `one.zero` II

- `method = "ratio"` and `one.zero = TRUE`

    ```{r, warning=FALSE, message=FALSE,  results="hide"}
    tab3 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
    varname = "langspan0913", method = "ratio", level = "tract", county = 1,
    tract = 950501, one.zero = TRUE)
    ```

    ```{r, echo = FALSE}
    tab3[, c(1,6, 11:12), with = FALSE]
    ```

    ```{r, eval = TRUE}
    ( sqrt(( 5.4711125 ^ 2 + 22.4924 ^ 2 )
    + ( 0.0226390685640362 ^ 2 * 102.1277 ^ 2 )) / 1546 ) * 1.645
    ```

---

##  `agg` and `one.zero` I

- `method = "agg"` and `one.zero = FALSE`

    ```{r,  results="hide"}
    tab4 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048)",
    varname = "test", method = "agg", level = "tract", county = 1,
    tract = 950501, one.zero = FALSE)
    ```

    ```{r, echo = FALSE}
      tab4[, c(1,6, 11:12), with = FALSE]
    ```

    ```{r, eval = TRUE}
    sqrt( 5.4711125 ^ 2 + 22.4924 ^ 2 + 5.4711125 ^ 2 )  * 1.645
    ```

---

## `agg` and `one.zero` II

- `method = "agg"` and `one.zero = TRUE`

    ```{r,  results="hide"}
    tab4 <- sumacs(formula = "(b16004_004 + b16004_026 + b16004_048)",
    varname = "test", method = "agg", level = "tract", county = 1,
    tract = 950501, one.zero = TRUE)
    ```

    ```{r, echo = FALSE}
    tab4[, c(1,6, 11:12), with = FALSE]
    ```

    ```{r, eval = TRUE}
    sqrt( 5.4711125 ^ 2 + 22.4924 ^ 2 )  * 1.645
    ```

---

## To note

- When the square root of the standard error formula doesn't exist (e.g., the square root of a negative number), the ratio formula is used  instead.

- The ratio adjustment is done `case by case`.

- There are some cases where the `one.zero` option makes the square root undefinable. In those cases, the function uses the "ratio" formula to compute standard errors.

- There is a possibility that the "ratio" estimates are higher than the "proportion" estimates without the `one.zero` option.

---

## Using an Excel file

- Save the Excel file as a CSV file after adjusting the format (e.g., no empty rows, name of variables)
- Read that file using `R`

    ```{r, echo = FALSE}
    path <- "/Users/sdaza/Documents/UWSociology/00Projects/Various/AppliedPopulationLab/"
    ```

    ```{r}
    sheet <- fread(paste0(path, "moe.csv"))
    str(sheet)
    ```

---

## Using an Excel file

- Define all levels in a vector

    ```{r}
    # create level vector
    levels <- c("state", "county", "county.subdivision",
    "tract", "block.group", "congressional.district",
    "school.district.secondary", "school.district.elementary")
    ```

- This can take a while! Here I downloaded the data first and then I saved them

    ```{r, eval = FALSE}
    # download the data first
    wi_acs <- acsdata(sheet[, formula], level = levels)
    # save the data
    save(wi_acs, file = "wi_acs.rd")
    ```

    ```{r, echo = FALSE}
    load(paste0(path, "wi_acs.rdata")) # load data
    ```


---

## Using an Excel file

- Define the function fields
- Open the output file in Excel

    ```{r}
    # compute indicators and export the data
    out <- sumacs(formula = sheet[, formula], varname = sheet[, myfield],
      method = sheet[, type], level = levels, one.zero = TRUE,
      data = wi_acs, file = "output.csv")
    ```

---

## Conclusions

- `Strengths`
    - Automatized and tailored way to obtain indicators
    - Reads formulas directly
    - Different formats for outputs (wide and long), exports the data
    - Easy way to adjust MOEs to different confidence levels
    - Ratio adjustment of standard errors case by case
    - Zero-option for proportions, ratios and aggregations

- `Limitations`
    - It depends heavily on the ACS package
    - Not the fastest function (loops involved despite some vectorization)
    - It takes less than a minute to create 90 indicators
    - Limited to only eight levels
**Last Update: `r format(Sys.time(), '%m/%d/%Y')`**

-----

### References

