---
layout: post
title: "Functions for sample size and error"
description: ""
category: survey
tags: [sampling, home]
---


Here two handy functions in R to compute sample sizes and errors for a proportion taking into account design effect (deff), response rate (rr), and finite population correction (fpc). Just there for those times when one needs to do these calculations quickly.

[Here](https://gist.github.com/sdaza/7896840) you can see my code.

```{r, echo=FALSE}
options(replace.assign=TRUE, tidy=TRUE, width=110, scipen = 4, digits = 3)
source("/Users/sdaza/Documents/Workplace/Functions/MyFunctions.R")
```


The inputs of these functions are:

- **n** = sample size
- **e** = sampling error
- **deff** = design effect, by default 1 (SRS)
- **rr** = response rate, by default 1
- **N** = population size, by default NULL (infinite population)
- **cl** = confidence level , by default .95
- **p** = proportion, by default 0.5 (maximum variance of a proportion)
- **relative** = to estimate relative error, by default FALSE

## serr: sampling error

An example for n = 400 and all the inputs at their default values:

```{r}
serr(400)
```

The output is rounded to 4 decimals. A more complete example:

- **n** = 400
- **deff** = 1.5
- **response rate** = 80%
- **population size** = 1000

```{r}
serr(n=400, deff=1.5, rr=.8, N=1000)
```

The sample size (n) has to be always lower than the population (N).  It is important to note that the final sample size to compute the sampling error comes from: $n = \frac{N}{deff} * rr$.

```{r}
serr(n=400, N=350)
```

## ssize: sample size

Let's get a sample size with an error of .03, a population of 1000 elements, a response rate of 0.80, and an effect design of 1.2:

```{r}
ssize(e=.03, deff=1.2, rr=.8, N=1000)
```

If the the sample size is bigger than the population because of low response rates or big design effects, the sample size will be fixed to N:

```{r}
ssize(e=.03, deff=5, rr=.6, N=1000)
```

## Working with strata

Finally, we can estimate different sample sizes by strata using vectors or a data frame:

```{r}
# example sampling frame (4 strata)
frame <- data.frame(
	strata = 1:4,
	N =c(10000, 5000, 2000, 1000),
	deff =c(1.1, 1, 1.3, .8),
	rr = c(.8, .9, .85,.8),
	p = c(.3, .6, .1, .2))
```

```{r, echo=FALSE}
frame
```

```{r}
frame$n1 <- ssize(e=.02, deff=frame$deff, rr=frame$rr, N=frame$N, p=frame$p)
frame$e1 <- serr(n=frame$n1, deff=frame$deff, rr=frame$rr, N=frame$N, p=frame$p)
```

```{r, echo=FALSE}
frame
```


## astrata: allocation of stratified sample

**astrata** is a function to define different allocations of a given total sample size across strata. So far, I have specified three options:

- Allocation by error: the sample size by strata is defined based on a given sampling error
- Square-root allocation: $\frac{sqrt(N)}{sum(sqrt(N)}
- Mixed: a combination between a proportional and equal-size allocation, where wp=1 means proportional, and wp=0 equal-size.


## serrst: sampling error with strata

The function **serrst** computes the sampling error of a stratified sample. We only have to define the sample of each stratum. Optionally we can specify the distribution of strata (N could be a number or proportion), and the proportion of the variable of interest (by default p=.05). Based on the example used above:

```{r}
serrst(n=frame$n1, N=frame$N, p=frame$p)
```



